---
title: "R Code - A general class of two-sample statistics for binary and  time-to-event outcomes"
author: "Marta Bofill Roig^[[marta.bofill.roig@upc.edu](marta.bofill.roig@upc.edu)], Guadalupe Gomez Melis^[[lupe.gomez@upc.edu](lupe.gomez@upc.edu)]"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    toc: true
    number_sections: false
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = F}
knitr::opts_chunk$set(
  collapse = T,
  comment = "#>"
)
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/marta.bofill/Dropbox/Marta-Lupe/2019 PAPER JRSS-B/GITHUB Code/Functions")
source('simulation-functions.R')
source('binary-functions.R')
source('survival-functions.R')
source("C:/Users/Marta.Bofill/Dropbox/Marta-Lupe/2019 PAPER JRSS-B/GITHUB Code/Functions/cov-functions-completekernel.R")
source("C:/Users/Marta.Bofill/Dropbox/Marta-Lupe/2019 PAPER JRSS-B/GITHUB Code/Functions/cov-functions-complete.R")
``` 


This report describes the R Code for the preprint *A general class of two-sample statistics for binary and  time-to-event outcomes*. Available at: [LINK-ARXIV](https://www.google.com)

## Prelude 

Package dependences:
```{r,warning=F}
library(survival)
library(zoo)
library(muhaz)
library(copula)
```

## Functions 

### L-statistics

Function `lstats()`: performs a L-statistic for right-censored and binary data.

Arguments:

  - time: The observed time.
  - status: The status indicator, normally 0=alive, 1=dead. 
  - binary: 0=nonresponse, 1=response. 
  - treat: The treatment-group indicator, normally 0=control, 1=intervention.
  - tau: follow-up. Default NULL denoting the last time in which both groups had patients at risk.
  - taub parameter: time-point at which binary endpoint is evaluated. If NULL, the binary endpoint is evaluated at tau.
  - rho: A scalar parameter that controls the type of test (see Details).
  - gamma: A scalar parameter that controls the type of test (see Details).
  - eta: A scalar parameter that controls the type of test (see Details).
  - gamma: A scalar parameter that controls the type of test (see Details).

Details:

Th L-statistics are defined as a weighted linear combination of the difference of proportions statistic for the binary outcome
and the integrated weighted difference of two survival functions for the time-to-event outcome, 
as follows, 
\begin{equation} \label{Lclass}
	%	\mathrm{\textbf{U}}_{n}^{\omega_b,\omega_s,\hat{Q}}(\tau_0,\tau_b,\tau)=
	\mathrm{\textbf{U}}_{n}^{\omega}(\tau_0,\tau_b,\tau; \hat{Q})
	=  {\omega}_b \cdot \frac{U_{b,n}\left(\tau_b\right)}{\hat{\sigma}_b}  
	+  {\omega}_s \cdot \frac{U_{s,n}(\tau_0,\tau;\hat{Q})}{\hat{\sigma}_s} 
\end{equation}

for some real numbers $\omega_b,\omega_s\in (0,1)$, such that  $\omega_b+\omega_s=1$, and where: 
\begin{eqnarray} 
	U_{b,n}\left(\tau_b\right) &=& \sqrt{\frac{n^{(0)}n^{(1)}}{n}} \left(\hat{p}^{(1)}\left(\tau_b\right)  -  \hat{p}^{(0)}\left(\tau_b\right)\right) \\ 
	U_{s,n}(\tau_0,\tau;\hat{Q}) &=& 
	\sqrt{\frac{n^{(0)}n^{(1)}}{n}}\left( \int_{\tau_0}^{\tau} \hat{Q}(t) \cdot\left( \hat{S}^{(1)}(t)-\hat{S}^{(0)}(t)\right)  dt \right) 
\end{eqnarray}
representing by $\hat{\sigma}_b^2$ and $\hat{\sigma}_s^2$ the estimated variances of $U_{b,n}\left(\tau_b\right)$ and $U_{s,n}(\tau_0,\tau;\hat{Q})$, respectively; and 
denoting by $\hat{p}^{(i)}(\tau_b)$ the estimated proportion of events $\varepsilon_b$ before  $\tau_b$, and by $\hat{S}^{(i)}(\cdot)$ the  Kaplan-Meier estimator of $S^{(i)}(\cdot)$. 

The term $\hat{Q}(\cdot)$  is a random function defined as:
\begin{equation}  
	\hat{Q}(t) = \hat{w}^{(i)}(t)^\eta \cdot (\hat{S}^{(i)}(t))^\rho \cdot (1-\hat{S}^{(i)}(t))^\gamma
\end{equation} 
for $\eta,\rho,\gamma\geq0$, and where:
$$\hat{w}(t) = \frac{n\hat{C}^{(0)}(t-)\hat{C}^{(1)}(t-)}{n^{(0)}\hat{C}^{(0)}(t-)+n^{(1)}\hat{C}^{(1)}(t-)}$$
is the weight proposed by Pepe-Fleming (Biometrics ,1989). 

More details can be found in the corresponding pre-print.

### Binary statistic 

Function `bintest()`: performs a test for testing the null hypothesis that the proportions (probabilities of success) in two groups are the same.

Arguments:

- binary: 0=nonresponse, 1=response. 
- treat: The treatment-group indicator, normally 0=control, 1=intervention.
- variance estimator: Unpooled/Pooled

Details:

The statistic used is:
\begin{eqnarray} 
	\frac{U_{b,n}\left(\tau_b\right)}{\hat{\sigma}_b} &=& \frac{\sqrt{\frac{n^{(0)}n^{(1)}}{n}} \left(\hat{p}^{(1)}\left(\tau_b\right)  -  \hat{p}^{(0)}\left(\tau_b\right)\right)}{\sqrt{\hat{p}(\tau_b)\cdot\left( 1-\hat{p}(\tau_b)\right)}} 
\end{eqnarray} 

More details can be found in the corresponding pre-print.

### Time-to-event statistic 

Function `survtest()`: performs a test for right-censored data. It uses the Weighted Kaplan-Meier family of statistics for testing the differences of two survival curves.

Arguments:

- time: The observed time.
- status: The status indicator, normally 0=alive, 1=dead. 
- treat: The treatment-group indicator, normally 0=control, 1=intervention.
- tau: follow-up. Default NULL denoting the last time in which both groups had patients at risk.
- rho: A scalar parameter that controls the type of test (see Weights).
- gamma: A scalar parameter that controls the type of test (see Weights).


Details:

The statistic used is:
\begin{eqnarray}  
	\frac{U_{s,n}(\tau_0,\tau;\hat{Q})}{\hat{\sigma}_s} &=& \frac{\sqrt{\frac{n^{(0)}n^{(1)}}{n}}\left( \int_{\tau_0}^{\tau} \hat{Q}(t) \cdot\left( \hat{S}^{(1)}(t)-\hat{S}^{(0)}(t)\right)  dt \right)}{\sqrt{-  \int_{\tau_0}^{\tau} \frac{\left(\int_{t}^{\tau} \hat{Q}(u)\hat{S}(u) du \right)^2 }{\hat{S}(t)\hat{S}(t-) }\cdot \frac{n^{(0)}\hat{C}^{(0)}(t-)+n^{(1)}\hat{C}^{(1)}(t-)}{n\hat{C}^{(0)}(t-)\hat{C}^{(1)}(t-)} d\hat{S}(t)}} 
\end{eqnarray} 
For more information on the weight function $\hat{Q}(t)$, see Details in  function `lstats()`.



### Covariance

Functions `survbinCov()` and `survbinCov_kernel()`: compute the covariance between the binary and time-to-event statistics. 

Arguments:

  - time: The observed time.
  - status: The status indicator, normally 0=alive, 1=dead. 
  - binary: 0=nonresponse, 1=response. 
  - treat: The treatment-group indicator, normally 0=control, 1=intervention.
  - tau: follow-up. Default NULL denoting the last time in which both groups had patients at risk.
  - taub parameter: time-point at which binary endpoint is evaluated. If NULL, the binary endpoint is evaluated at tau.
  - rho: A scalar parameter that controls the type of test (see Details in function `lstats()`).
  - gamma: A scalar parameter that controls the type of test (see Details in function `lstats()`).
  - eta: A scalar parameter that controls the type of test (see Details in function `lstats()`).
  - gamma: A scalar parameter that controls the type of test (see Details in function `lstats()`).

Details:

These functions return an estimation of $\mathrm{Cov}(U_{b,n}, U_{s,n}(Q))$ given by:

\begin{eqnarray*}  
\hat{Cov} &=& \sum_{i=0,1} \frac{n-n^{(i)}}{n} \left( 
    \int_{\tau_0}^{\tau_b}   \frac{\int_{s}^{\tau} \hat{Q}(t) \hat{S}^{(i)}(t)dt}{\hat{S}^{(i)}(s)} \cdot \left(\hat{p}_N^{(i)}(s)-\hat{p}^{(i)}  \right) d\hat{S}^{(i)}(s) \right. \\
    && \left. + \int_{\tau_b}^{\tau}
    \frac{\hat{p}^{(i)} \cdot \int_{s}^{\tau} \hat{Q}(t) \hat{S}^{(i)}(t)dt }{\hat{S}^{(i)}(s) }\left( d\hat{S}_X^{(i)}(s) -d\hat{S}^{(i)}(s)\right) \right)  
\end{eqnarray*}

The difference between these two functions is that we use different methods for estimating $p_N^{(i)}(s)$:

  - Function `survbinCov()`: The estimation is made by using the plug-in estimation:
  \begin{eqnarray*}  
  		\hat{p}_N^{(i)}(t) &=& \displaystyle \sum_{j=1,...,n^{(i)}} \frac{ I\{X_{ij}=1, dN_{ij}(t)=1 \}}{d\bar{N}^{(i)}(t)} =  \frac{\sum_{j=1,...,n^{(i)}} I\{X_{ij}=1, dN_{ij}(t)=1 \}}{\sum_{j=1,...,n^{(i)}} I\{dN_{ij}(t)=1 \}}
  	\end{eqnarray*} 
  	
- Function `survbinCov_kernel()`: The estimation is made by using smoothing estimating methods. We first notice that:
  \begin{eqnarray*}  
  		\mathrm{E}(X_{ij} dN_{j}^{(i)}(t)) &=& \mathrm{E}(X_{ij}Y_{j}^{(i)}(t)\lambda_X^{(i)}(t)dt) = \mathrm{E}(X_{ij}Y_{j}^{(i)}(t))\lambda_X^{(i)}(t)dt 
  	\end{eqnarray*} 
Then:
\begin{eqnarray*}  
  		p_N^{(i)}(t)=P(X_{ij}=1| dN_{ij}(t)=1) &=& \frac{\mathrm{P}(X_{ij} \cdot dN_{ij}(t)=1)}{\mathrm{P}(dN_{ij}(t)=1)} = \frac{\mathrm{E}(X_{ij}Y_{j}^{(i)}(t))\lambda_X^{(i)}(t)dt }{\mathrm{E}(Y_{j}^{(i)}(t))\lambda^{(i)}(t)dt } \\
  		&=& \frac{\mathrm{P}(X_{ij}Y_{j}^{(i)}(t)=1)\lambda_X^{(i)}(t)dt }{\mathrm{P}(Y_{j}^{(i)}(t)=1)\lambda^{(i)}(t)dt } 
\end{eqnarray*}
We estimate the last expression as:
\begin{eqnarray*}  
  		\hat{p}_N^{(i)}(t)=  \frac{\sum_{j=1,...,n^{(i)}}I\{X_{ij}Y_{j}^{(i)}(t)=1\}\hat{\lambda}_X^{(i)}(t)  }{\sum_{j=1,...,n^{(i)}}I\{Y_{j}^{(i)}(t)=1\}\hat{\lambda}^{(i)}(t)  } 
\end{eqnarray*}
where the hazard functions $\hat{\lambda}_X^{(i)}(t)$ and  $\hat{\lambda}^{(i)}(t)$ are estimated by kernel smoothing methods (`R` package `muhaz`).

More details can be found in the corresponding pre-print.

## Checking the functions 

For each function `bintest()`, `survtest()` and `lstats()`, we simulate $nsim$ trials to check the null distribution of each statistic and to evaluate the significance level $\alpha$ that we would get from each statistic.

The considered $\alpha$ and $nsim=1000$ are the following: 
```{r}
alpha=0.025;
z.alpha <- qnorm(1-alpha,0,1) 
nsim=1000
```
Note that simulating $nsim=1000$ trials, we obtain the following CI for the estimated $\alpha$:
```{r} 
sd=sqrt(alpha*(1-alpha)/nsim)
c(alpha-z.alpha*sd,alpha+z.alpha*sd) 
```


### Testing the function `lstats()` (L-statistics statistic) 

Simulation parameters:
```{r}
p=0.1
set.seed(43321)
nsim=1000
```
Varying the association parameter of the copula, we obtain  the following Spearman's rho for the bivariate copula:
```{r}
rho(frankCopula(param=1,dim=2)) 
rho(frankCopula(param=2,dim=2)) 
rho(frankCopula(param=5,dim=2)) 
rho(frankCopula(param=10,dim=2)) 
```

Parameters for the survival and binary distributions:
```{r}
a.shape=1.5 # shape parameter weibull distr
b.scale=2 # scale parameter weibull distr
rate.param=0.5 # rate parameter exp distribution (censoring distribution)
prob0=0.4 # probability binary outcome
ass.par=c(2,5) # association parameter copula
ss=500 # sample size per arm
```
Parameters of the problem:
```{r}
tau0=0
tau=3 # end of follow-up survival
taub=c(1,3) # follow-up binary outcome
rho=0 # parameters for the weights, see details in lstats() function
gam=0 # parameters for the weights, see details in lstats() function
eta=1 # parameters for the weights, see details in lstats() function
```

We simulate four different scenarios, varying the association parameter and $\tau_b$. We evaluate the performance of the statistics using both the plug-in covariance estimation and the smoothing estimation.

Simulated cases:

  - Case 1: $\tau_b=1$, association parameter copula$=2$
  - Case 2: $\tau_b=1$, association parameter copula$=5$
  - Case 3: $\tau_b=3$, association parameter copula$=2$
  - Case 4: $\tau_b=3$, association parameter copula$=5$


For all the scenarios, we study the distribution of the L-statistics under $H_{0,s}=H_{0,b}\cap H_{0,s}$, so that the simulated data is such that the distributions are the same  in the control and treatment group. 

#### Using plug-in

Initial timing and seed for the simulations.
```{r}
t0=Sys.time()
set.seed(9110)
```

Code for simulating bivariate binary and survival data is not displayed.
<!-- case 1 -->
```{r, echo=F}
zb_cs1 <- c()
ub <- c()
ub_sd <- c()
zs_cs1 <- c()
us <- c()
us_sd <- c()
xcov <- c() 
zsb_cs1 <- c()

for(i in 1:nsim){
  data=simsurvbin(a.shape=a.shape, b.scale=b.scale, rate.param=rate.param, prob0=prob0, 
           ass.par=2, #ass.par=ass.par[1], 
           ss=ss)
  
  aux <- bintest(data$binary, data$treat)
  zb_cs1[i] <- aux[1]
  ub[i] <- aux[2]
  ub_sd[i] <- aux[3]
  
  aux <- survtest(data$time, data$status, data$treat, tau=tau, rho=rho, gam=gam, eta=eta) 
  zs_cs1[i] <- aux[1]
  us[i] <- aux[2]
  us_sd[i] <- aux[3]
  
  xcov[i] <- survbinCov(data$time, data$status, data$binary, data$treat, tau0=0, tau=tau, 
                        taub= 1, #taub=taub[1], 
                        rho=rho, gam=gam, eta=eta) 
  
  zsb_cs1[i] <- (zb_cs1[i]+zs_cs1[i])/sqrt(2+2*xcov[i]/(ub_sd[i]*us_sd[i]))
} 
```
<!-- case 2 -->
```{r, echo=F}
zb_cs2 <- c()
ub <- c()
ub_sd <- c()
zs_cs2 <- c()
us <- c()
us_sd <- c()
xcov <- c() 
zsb_cs2 <- c()

for(i in 1:nsim){
  data=simsurvbin(a.shape=a.shape, b.scale=b.scale, rate.param=rate.param, prob0=prob0, 
           ass.par=5, #ass.par=ass.par[1], 
           ss=ss)
  
  aux <- bintest(data$binary, data$treat)
  zb_cs2[i] <- aux[1]
  ub[i] <- aux[2]
  ub_sd[i] <- aux[3]
  
  aux <- survtest(data$time, data$status, data$treat, tau=tau, rho=rho, gam=gam, eta=eta) 
  zs_cs2[i] <- aux[1]
  us[i] <- aux[2]
  us_sd[i] <- aux[3]
  
  xcov[i] <- survbinCov(data$time, data$status, data$binary, data$treat, tau0=0, tau=tau, 
                        taub=1,#taub=taub[1], 
                        rho=rho, gam=gam, eta=eta) 
  
  zsb_cs2[i] <- (zb_cs2[i]+zs_cs2[i])/sqrt(2+2*xcov[i]/(ub_sd[i]*us_sd[i]))
} 
```
<!-- case 3 -->
```{r, echo=F}
zb_cs3 <- c()
ub <- c()
ub_sd <- c()
zs_cs3 <- c()
us <- c()
us_sd <- c()
xcov <- c() 
zsb_cs3 <- c()

for(i in 1:nsim){
  data=simsurvbin(a.shape=a.shape, b.scale=b.scale, rate.param=rate.param, prob0=prob0, 
           ass.par=2, #ass.par=ass.par[1], 
           ss=ss)
  
  aux <- bintest(data$binary, data$treat)
  zb_cs3[i] <- aux[1]
  ub[i] <- aux[2]
  ub_sd[i] <- aux[3]
  
  aux <- survtest(data$time, data$status, data$treat, tau=tau, rho=rho, gam=gam, eta=eta) 
  zs_cs3[i] <- aux[1]
  us[i] <- aux[2]
  us_sd[i] <- aux[3]
  
  xcov[i] <- survbinCov(data$time, data$status, data$binary, data$treat, tau0=0, tau=tau, 
                        taub=3, #taub=taub[2], 
                        rho=rho, gam=gam, eta=eta) 
  
  zsb_cs3[i] <- (zb_cs3[i]+zs_cs3[i])/sqrt(2+2*xcov[i]/(ub_sd[i]*us_sd[i]))
} 
```
<!-- case 4 -->
```{r, echo=F}
zb_cs4 <- c()
ub <- c()
ub_sd <- c()
zs_cs4 <- c()
us <- c()
us_sd <- c()
xcov <- c() 
zsb_cs4 <- c()

for(i in 1:nsim){
  data=simsurvbin(a.shape=a.shape, b.scale=b.scale, rate.param=rate.param, prob0=prob0, 
           ass.par=5, #ass.par=ass.par[2], 
           ss=ss)
  
  aux <- bintest(data$binary, data$treat)
  zb_cs4[i] <- aux[1]
  ub[i] <- aux[2]
  ub_sd[i] <- aux[3]
  
  aux <- survtest(data$time, data$status, data$treat, tau=tau, rho=rho, gam=gam, eta=eta) 
  zs_cs4[i] <- aux[1]
  us[i] <- aux[2]
  us_sd[i] <- aux[3]
  
  xcov[i] <- survbinCov(data$time, data$status, data$binary, data$treat, tau0=0, tau=tau, 
                        taub=3, #taub=taub[2], 
                        rho=rho, gam=gam, eta=eta) 
  
  zsb_cs4[i] <- (zb_cs4[i]+zs_cs4[i])/sqrt(2+2*xcov[i]/(ub_sd[i]*us_sd[i]))
} 
```

**Distribution of the L-statistic under the null hypothesis:**

Plot distribution: blue densities for the binary statistic and survival statistic; red density for the N(0,1); black density for the binary and survival statistic (L-statistic)

```{r, fig.height=8, fig.width=8, echo=F}
x=rnorm(n=10000)

par(mfrow = c(2, 2))

plot(density(zsb_cs1, na.rm = T),ylim=c(0,0.6), lwd = 2, main='Density Case 1') 
lines(density(zb_cs1), col=4)
lines(density(zs_cs1), col=4)
lines(density(x), col=2)

plot(density(zsb_cs2, na.rm = T),ylim=c(0,0.6), lwd = 2, main='Density Case 2') 
lines(density(zb_cs2), col=4)
lines(density(zs_cs2), col=4) 
lines(density(x), col=2)

plot(density(zsb_cs3, na.rm = T),ylim=c(0,0.6), lwd = 2, main='Density Case 3')  
lines(density(zb_cs3), col=4)
lines(density(zs_cs3), col=4)  
lines(density(x), col=2)

plot(density(zsb_cs4, na.rm = T),ylim=c(0,0.6), lwd = 2, main='Density Case 4') 
lines(density(zb_cs4), col=4)
lines(density(zs_cs4), col=4) 
lines(density(x), col=2)
```


Mean and Standard deviation: 
```{r}
summary(zsb_cs1)
mean(zsb_cs1, na.rm = T); sd(zsb_cs1, na.rm = T) 
summary(zsb_cs2)
mean(zsb_cs2, na.rm = T); sd(zsb_cs2, na.rm = T) 
summary(zsb_cs3)
mean(zsb_cs3, na.rm = T); sd(zsb_cs3, na.rm = T) 
summary(zsb_cs4)
mean(zsb_cs4, na.rm = T); sd(zsb_cs4, na.rm = T) 
```


Timing:
```{r}
(t1=Sys.time()-t0)
```

**Estimated $\alpha$**
```{r}
sum(zsb_cs1> z.alpha, na.rm = T)/nsim 
sum(zsb_cs2> z.alpha, na.rm = T)/nsim 
sum(zsb_cs3> z.alpha, na.rm = T)/nsim 
sum(zsb_cs4> z.alpha, na.rm = T)/nsim 
```


#### Using  kernel smoothing estimation

Initial timing and seed for the simulations.
```{r}
t0=Sys.time()
set.seed(9110)
```

Code for simulating bivariate binary and survival data is not displayed.
<!-- case 1 -->
```{r, echo=F}
zb_cs1k <- c()
ub <- c()
ub_sd <- c()
zs_cs1k <- c()
us <- c()
us_sd <- c()
xcov_kernel <- c() 
zsb_cs1k <- c()

for(i in 1:nsim){
  data=simsurvbin(a.shape=a.shape, b.scale=b.scale, rate.param=rate.param, prob0=prob0, 
           ass.par=2, #ass.par=ass.par[1], 
           ss=ss)
  
  aux <- bintest(data$binary, data$treat)
  zb_cs1k[i] <- aux[1]
  ub[i] <- aux[2]
  ub_sd[i] <- aux[3]
  
  aux <- survtest(data$time, data$status, data$treat, tau=tau, rho=rho, gam=gam, eta=eta) 
  zs_cs1k[i] <- aux[1]
  us[i] <- aux[2]
  us_sd[i] <- aux[3]
  
  xcov_kernel[i] <- survbinCov_kernel(data$time, data$status, data$binary, data$treat, tau0=0, tau=tau, 
                                      taub=1, #taub=taub[1], 
                                      rho=rho, gam=gam, eta=eta) 
  
  zsb_cs1k[i] <- (zb_cs1k[i]+zs_cs1k[i])/sqrt(2+2*xcov_kernel[i]/(ub_sd[i]*us_sd[i]))
} 
```
<!-- case 2 -->
```{r, echo=F}
zb_cs2k <- c()
ub <- c()
ub_sd <- c()
zs_cs2k <- c()
us <- c()
us_sd <- c()
xcov_kernel <- c() 
zsb_cs2k <- c()

for(i in 1:nsim){
  data=simsurvbin(a.shape=a.shape, b.scale=b.scale, rate.param=rate.param, prob0=prob0, 
           ass.par=5, #ass.par=ass.par[1], 
           ss=ss)
  
  aux <- bintest(data$binary, data$treat)
  zb_cs2k[i] <- aux[1]
  ub[i] <- aux[2]
  ub_sd[i] <- aux[3]
  
  aux <- survtest(data$time, data$status, data$treat, tau=tau, rho=rho, gam=gam, eta=eta) 
  zs_cs2k[i] <- aux[1]
  us[i] <- aux[2]
  us_sd[i] <- aux[3]
  
  xcov_kernel[i] <- survbinCov_kernel(data$time, data$status, data$binary, data$treat, tau0=0, tau=tau, 
                                      taub=1,#taub=taub[1], 
                                      rho=rho, gam=gam, eta=eta) 
  
  zsb_cs2k[i] <- (zb_cs2k[i]+zs_cs2k[i])/sqrt(2+2*xcov_kernel[i]/(ub_sd[i]*us_sd[i]))
} 
```
<!-- case 3 -->
```{r, echo=F}
zb_cs3k <- c()
ub <- c()
ub_sd <- c()
zs_cs3k <- c()
us <- c()
us_sd <- c()
xcov <- c() 
zsb_cs3k <- c()

for(i in 1:nsim){
  data=simsurvbin(a.shape=a.shape, b.scale=b.scale, rate.param=rate.param, prob0=prob0, 
           ass.par=2, #ass.par=ass.par[1], 
           ss=ss)
  
  aux <- bintest(data$binary, data$treat)
  zb_cs3k[i] <- aux[1]
  ub[i] <- aux[2]
  ub_sd[i] <- aux[3]
  
  aux <- survtest(data$time, data$status, data$treat, tau=tau, rho=rho, gam=gam, eta=eta) 
  zs_cs3k[i] <- aux[1]
  us[i] <- aux[2]
  us_sd[i] <- aux[3]
  
  xcov_kernel[i] <- survbinCov_kernel(data$time, data$status, data$binary, data$treat, tau0=0, tau=tau, 
                                      taub=3, #taub=taub[2], 
                                      rho=rho, gam=gam, eta=eta) 
  
  zsb_cs3k[i] <- (zb_cs3k[i]+zs_cs3k[i])/sqrt(2+2*xcov_kernel[i]/(ub_sd[i]*us_sd[i]))
} 
```
<!-- case 4 -->
```{r, echo=F}
zb_cs4k <- c()
ub <- c()
ub_sd <- c()
zs_cs4k <- c()
us <- c()
us_sd <- c()
xcov_kernel <- c() 
zsb_cs4k <- c()

for(i in 1:nsim){
  data=simsurvbin(a.shape=a.shape, b.scale=b.scale, rate.param=rate.param, prob0=prob0, 
           ass.par=5, #ass.par=ass.par[2], 
           ss=ss)
  
  aux <- bintest(data$binary, data$treat)
  zb_cs4k[i] <- aux[1]
  ub[i] <- aux[2]
  ub_sd[i] <- aux[3]
  
  aux <- survtest(data$time, data$status, data$treat, tau=tau, rho=rho, gam=gam, eta=eta) 
  zs_cs4k[i] <- aux[1]
  us[i] <- aux[2]
  us_sd[i] <- aux[3]
  
  xcov_kernel[i] <- survbinCov_kernel(data$time, data$status, data$binary, data$treat, tau0=0, tau=tau, 
                                      taub=3,  #taub=taub[2], 
                                      rho=rho, gam=gam, eta=eta) 
  
  zsb_cs4k[i] <- (zb_cs4k[i]+zs_cs4k[i])/sqrt(2+2*xcov_kernel[i]/(ub_sd[i]*us_sd[i]))
} 
```

**Distribution of the L-statistic under the null hypothesis:**

Plot distribution: blue densities for the binary statistic and survival statistic; red density for the N(0,1); black density for the binary and survival statistic (L-statistic) 
```{r, fig.height=8, fig.width=8, echo=F}
x=rnorm(n=10000) 

par(mfrow = c(2, 2))
plot(density(zsb_cs1k, na.rm = T),ylim=c(0,0.6), lwd = 2, main='Density  Case 1') 
lines(density(zb_cs1k), col=4)
lines(density(zs_cs1k), col=4)
lines(density(x), col=2)

plot(density(zsb_cs2k, na.rm = T),ylim=c(0,0.6), lwd = 2, main='Density  Case 2') 
lines(density(zb_cs2k), col=4)
lines(density(zs_cs2k), col=4)
lines(density(x), col=2)

plot(density(zsb_cs3k, na.rm = T),ylim=c(0,0.6), lwd = 2, main='Density Case 3') 
lines(density(zb_cs3k), col=4)
lines(density(zs_cs3k), col=4)
lines(density(x), col=2)

plot(density(zsb_cs4k, na.rm = T),ylim=c(0,0.6), lwd = 2, main='Density  Case 4') 
lines(density(zb_cs4k), col=4)
lines(density(zs_cs4k), col=4)
lines(density(x), col=2)
```

Mean and Standard deviation: 
```{r}
summary(zsb_cs1k)
mean(zsb_cs1k, na.rm = T); sd(zsb_cs1k, na.rm = T) 
summary(zsb_cs2k)
mean(zsb_cs2k, na.rm = T); sd(zsb_cs2k, na.rm = T) 
summary(zsb_cs3k)
mean(zsb_cs3k, na.rm = T); sd(zsb_cs3k, na.rm = T) 
summary(zsb_cs4k)
mean(zsb_cs4k, na.rm = T); sd(zsb_cs4k, na.rm = T) 
```


Timing:
```{r}
(t1=Sys.time()-t0)
```

**Estimated $\alpha$**
```{r}
sum(zsb_cs1k> z.alpha, na.rm = T)/nsim 
sum(zsb_cs2k> z.alpha, na.rm = T)/nsim 
sum(zsb_cs3k> z.alpha, na.rm = T)/nsim 
sum(zsb_cs4k> z.alpha, na.rm = T)/nsim 
```


#### Comparison covariance estimation under different methods

The first is calculated by the usual cov funcion in R, the second is calculated by plug-in, the third is calculated by kernel-methods.
```{r}
cov(ub,us)
summary(xcov) 
summary(xcov_kernel) 
```


### Testing the function `bintest()` (Binary statistic)

Simulation parameters:
```{r, eval=T}
p=0.1
ss=500
set.seed(43321)
nsim=1000
```

**Distribution of the binary statistic under the null hypothesis $H_{0,b}$:** 

Code for simulating $nsim$ binary statistics under $H_{0,b}$ (not displayed).
```{r,echo=F, eval=T}
sim_bintest <- c()
sim_bintestp <- c()
for(i in 1:nsim){
  u1 <- runif(n=ss)
  u0 <- runif(n=ss)
  BE1 = ifelse(u1<p, 1, 0)
  treat1 = rep(1,ss)
  BE0 = ifelse(u0<p, 1, 0)
  treat0 = rep(0,ss)
  data = data.frame(binary=c(BE1, BE0), treat=c(treat1,treat0))
  
  sim_bintest[i] = bintest(data$binary, data$treat)[1]
}
for(i in 1:nsim){
  p=0.1
  u1 <- runif(n=ss)
  u0 <- runif(n=ss)
  BE1 = ifelse(u1<p, 1, 0)
  treat1 = rep(1,ss)
  BE0 = ifelse(u0<p, 1, 0)
  treat0 = rep(0,ss)
  data = data.frame(binary=c(BE1, BE0), treat=c(treat1,treat0))
  
  sim_bintestp[i] = bintest(data$binary, data$treat, x3="Pooled")[1]
}
```
Distribution:
```{r, fig.width=5, fig.height=5, echo= F, eval=T}
plot(density(sim_bintest), ylim=c(0,0.6), main= 'Distribution binary test')
lines(density(sim_bintestp))
x=rnorm(n=100000) 
lines(density(x), col=2)
```

```{r, eval=T}
mean(sim_bintestp)
sd(sim_bintestp)
```

**Estimated $\alpha$**

We estimate $\alpha$ using the unpooled variance estimator (`sim_bintest`) and using pooled variance estimator  (`sim_bintestp`):
```{r, eval=T}
sum(sim_bintest> z.alpha)/nsim
sum(sim_bintestp> z.alpha)/nsim 
```


### Testing the function `survtest()` (Time-to-event statistic)

**Distribution of the survival statistic under the null hypothesis $H_{0,s}$:**  

Simulation parameters:
```{r, eval=T} 
shape_param=0.5; scale_param=1; rate_param=1;
ss = 500
nsim= 1000
set.seed(123321)
```

Code for simulating $nsim$ time-to-event statistics under $H_{0,s}$ (not displayed).
```{r, echo= F, eval=T} 
t0=Sys.time()
sim_survtest <- c()

for(i in 1:nsim){
  u1 <- runif(n=ss)
  u0 <- runif(n=ss)

  TE1 = scale_param*(-log(1-u1))^(1/shape_param)
  TC1 = rexp(n= ss, rate = rate_param)
  time1= ifelse(TE1<=TC1, TE1,TC1)
  status1 = ifelse(TE1<=TC1,1,0)
  treat1 = rep(1,ss)
  TE0 = scale_param*(-log(1-u0))^(1/shape_param)
  TC0 = rexp(n=ss, rate = rate_param)
  time0= ifelse(TE0<=TC0, TE0,TC0)
  status0 = ifelse(TE0<=TC0,1,0)
  treat0 = rep(0,ss)

  data = data.frame(time=c(time1,time0), status=c(status1,status0),treat=c(treat1,treat0))
  sim_survtest[i] = survtest(data$time, data$status, data$treat, rho=0, gam=0, tau=3)[1] 
} 
t1=Sys.time()-t0
(t1) 
```
Summary distribution statistic:
```{r,eval=T}
summary(sim_survtest)
```

```{r, fig.width=5, fig.height=5, echo= F, eval=T}
x=rnorm(n=100000)
plot(density(sim_survtest), ylim=c(0,0.6), main='Distribution time-to-event test')
lines(density(x), col=2)
```

**Estimated $\alpha$**

Based on the previous $nsim$ trials, we estimated the $\alpha$ level, obtaining: 
```{r, eval=T}
sum(sim_survtest> z.alpha)/nsim
```


 